---
title: "simulation_study"
format: html
editor: visual
---

# Simulation studies

## Simulation

For now, we focus on the case of **Bullseye**.

Correctly specified prior.

```{r}

library(MASS)

# Dartboard region classifier
get_dartboard_region_vectorized <- function(x, y) {
  segment_numbers <- c(20, 1, 18, 4, 13, 6, 10, 15, 2, 17,
                       3, 19, 7, 16, 8, 11, 14, 9, 12, 5)
  segment_width <- 360 / 20
  
  bull_inner <- 6.35
  bull_outer <- 15.9
  triple_inner <- 99
  triple_outer <- 107
  double_inner <- 162
  double_outer <- 170

  r <- sqrt(x^2 + y^2)
  theta <- (90 - atan2(y, x) * 180 / pi) %% 360
  segment_idx <- floor(theta / segment_width) + 1
  segment <- segment_numbers[segment_idx]

  region <- character(length(x))
  region[r <= bull_inner] <- "Bull"
  region[r > bull_inner & r <= bull_outer] <- "Outer Bull"
  region[r > bull_outer & r <= triple_inner] <- paste0("S", segment[r > bull_outer & r <= triple_inner])
  region[r > triple_inner & r <= triple_outer] <- paste0("T", segment[r > triple_inner & r <= triple_outer])
  region[r > triple_outer & r <= double_inner] <- paste0("S", segment[r > triple_outer & r <= double_inner])
  region[r > double_inner & r <= double_outer] <- paste0("D", segment[r > double_inner & r <= double_outer])
  region[r > double_outer] <- "Miss"

  return(region)
}

# Simulate data from prior G
set.seed(123)
n_players <- 5
throws_per_player <- 10

lognormal_mean <- 5    # set mean of log(sigma^2)
lognormal_sd <- 0.5    # set sd of log(sigma^2)

# Sample sigma_i^2 from LogNormal
sigma2_i <- rlnorm(n_players, meanlog = lognormal_mean, sdlog = lognormal_sd)
sigma_i <- sqrt(sigma2_i)  # Standard deviations

# Simulate dart throws for each player
player_data <- list()

for (i in 1:n_players) {
  samples <- MASS::mvrnorm(throws_per_player, mu = c(0, 0), Sigma = diag(sigma_i[i]^2, 2))
  
  # Debugging: Show (x, y) coordinates
  cat(sprintf("Player %d (sigma = %.2f) coordinates:\n", i, sigma_i[i]))
  print(samples)
  
  regions <- get_dartboard_region_vectorized(samples[,1], samples[,2])
  player_data[[i]] <- list(samples = samples, regions = regions)
  
  # Region counts for player
  cat(sprintf("Player %d region counts:\n", i))
  print(table(regions))
  cat("\n")
}

# Optional: Combine all players' regions
all_regions <- unlist(lapply(player_data, function(p) p$regions))
overall_counts <- table(all_regions)
cat("Overall region counts:\n")
print(overall_counts)

```

Then, we fit our EB model.

```{r}
library(stats4)
#fit our EB

#----convert this into Annuli data----

convert_regions_to_annuli <- function(region_counts) {
  annuli_counts <- rep(0, 6)  # Annuli 1 to 6
  
  for (region in names(region_counts)) {
    count <- region_counts[[region]]
    
    if (region == "Bull") {
      annuli_counts[1] <- annuli_counts[1] + count
    } else if (region == "Outer Bull") {
      annuli_counts[2] <- annuli_counts[2] + count
    } else if (startsWith(region, "T")) {
      annuli_counts[4] <- annuli_counts[4] + count
    } else if (startsWith(region, "D")) {
      annuli_counts[6] <- annuli_counts[6] + count
    } else if (startsWith(region, "S")) {
      # Single regions
      annuli_counts[3] <- annuli_counts[3] + count
    } else if (region == "Miss") {
      # Optionally ignore or lump separately
      next
    }
  }
  
  return(annuli_counts)
}

# Example usage
# Assume overall_counts is your 'table' from the previous step
annuli_counts <- convert_regions_to_annuli(overall_counts)
names(annuli_counts) <- paste0("Annulus_", 1:6)
print(annuli_counts)


#----compute the MLE prior----

# Radii for Bullseye annuli (mm)
bull_radii <- c(0, 6.35, 15.9, 99, 107, 162, 170)

# Exact annulus probability given a sigma^2
dart_annulus_probability <- function(r1, r2, sigma2) {
  pchisq(r2^2 / sigma2, df = 2) - pchisq(r1^2 / sigma2, df = 2)
}

# Log-likelihood function
log_likelihood_bull_ab <- function(a, b, observed_counts, radii = bull_radii, n_theta_samples = 5000) {
  if (a <= 0 || b <= 0) return(-Inf)

  # Sample theta from inverse-chi-squared(a, b)
  theta_samples <- b / rchisq(n_theta_samples, df = a)
  logL_total <- 0

  for (i in seq_along(observed_counts)) {
    n_obs <- observed_counts[i]
    if (n_obs == 0) next

    probs <- dart_annulus_probability(radii[i], radii[i+1], theta_samples)
    avg_prob <- mean(probs)
    avg_prob <- max(avg_prob, 1e-12)  # Numerical stability

    logL_total <- logL_total + n_obs * log(avg_prob)
  }
  
  return(logL_total)
}

# MLE wrapper
mle_bull_ab <- function(observed_counts, radii = bull_radii) {
  neg_log_likelihood <- function(a, b) {
    -log_likelihood_bull_ab(a, b, observed_counts, radii)
  }
  
  mle_result <- mle(neg_log_likelihood, start = list(a = 5, b = 1000), method = "L-BFGS-B", lower = c(0.1, 0.1))
  return(mle_result)
}

# --- Example usage ---

# Suppose your annuli counts are like this (after conversion):
observed_counts_bull <- annuli_counts

# Estimate (a, b)
set.seed(123)
mle_result <- mle_bull_ab(observed_counts_bull)
summary(mle_result)



#----compute the likelihoods----
# Exact annulus probability for given sigma^2
dart_annulus_probability <- function(r1, r2, sigma2) {
  pchisq(r2^2 / sigma2, df = 2) - pchisq(r1^2 / sigma2, df = 2)
}

# Function to compute LOG-likelihood first, then exp
log_likelihood_bull_given_sigma <- function(sigma2, observed_counts, radii = bull_radii) {
  log_likelihood <- 0
  
  for (i in seq_along(observed_counts)) {
    n_obs <- observed_counts[i]
    if (n_obs == 0) next

    prob_annulus <- dart_annulus_probability(radii[i], radii[i+1], sigma2)
    prob_annulus <- max(prob_annulus, 1e-12)  # Numerical stability to avoid log(0)

    log_likelihood <- log_likelihood + n_obs * log(prob_annulus)
  }
  
  return(log_likelihood)
}

# --- Example usage ---

# Suppose your annuli counts
observed_counts_bull <- c(2, 6, 2, 0, 0, 0)  # (n_1, ..., n_6)

sigma_vals <- seq(2, 25, length.out = 100)
sigma2_vals <- sigma_vals^2

log_likelihood_vals <- sapply(sigma2_vals, function(s2) {
  log_likelihood_bull_given_sigma(s2, observed_counts_bull)
}, USE.NAMES = FALSE)

likelihood_vals <- exp(log_likelihood_vals)
likelihood_vals <- likelihood_vals / max(likelihood_vals)  # Optional normalization

plot(sigma2_vals, likelihood_vals)

#----multiply to get the posterior----
mle_a <- coef(mle_result_bull)[["a"]]
mle_b <- coef(mle_result_bull)[["b"]]
prior <- dinvchisq(sigma2_vals, mle_a, mle_b)
unnormalized_posterior <- likelihood_vals * prior
normalized_posterior <- unnormalized_posterior / trapz(sigma2_vals, unnormalized_posterior)
plot(sigma2_vals, normalized_posterior)



#---calculate MAP ------
idx_max <- which.max(normalized_posterior)
sigma2_map <- sigma2_vals[idx_max]
cat("Posterior mode (MAP estimate of sigma^2):", sigma2_map, "\n")

```

Next, we fit HW EB model.

(See simulation_study_risk.qmd)

## Heat-map generation

-   Give some context for what it means for $\sigma = 300$ or something of that sort. Otherwise, it exists in a vacuum and is meaningless.

Here is a sample heat map.

```{r}

library(ggplot2)
library(ggforce)

# Parameters
mu <- c(0, 0)
sigma2 <- 3000  # variance, not sigma
sigma <- sqrt(sigma2)

# Dartboard ring radii (mm)
bullseye_inner <- 6.35
bullseye_outer <- 15.9
triple_inner <- 99
triple_outer <- 107
double_inner <- 162
double_outer <- 170

# Define grid
grid_lim <- 200  # display range
res <- 400       # resolution
x_seq <- seq(-grid_lim, grid_lim, length.out = res)
y_seq <- x_seq
grid <- expand.grid(x = x_seq, y = y_seq)

# Compute radial distance and density
grid$r <- sqrt(grid$x^2 + grid$y^2)
grid$density <- (1 / (2 * pi * sigma2)) * exp(-grid$r^2 / (2 * sigma2))

# Dartboard rings (for visual overlay)
rings <- data.frame(
  x0 = 0,
  y0 = 0,
  r = c(bullseye_inner, bullseye_outer, triple_inner, triple_outer, double_inner, double_outer)
)

# Plot symmetric heatmap from closed-form
ggplot(grid, aes(x = x, y = y)) +
  geom_raster(aes(fill = density)) +
  scale_fill_viridis_c(option = "plasma", trans = "sqrt") +
  geom_circle(data = rings, aes(x0 = x0, y0 = y0, r = r),
              color = "white", linetype = "dashed", inherit.aes = FALSE) +
  coord_fixed() +
  labs(title = expression("Heatmap of Bivariate Normal Density (N(0,"~sigma^2~"I))"),
       x = "X-coordinate (mm)", y = "Y-coordinate (mm)", fill = "Density") +
  theme_minimal()




```

Now, we want to plot our choice of $\sigma^2$ against the likelihood of actually landing in Bullseye.

```{r}
# Radius of the bullseye in mm
r_bull <- 6.35

# Grid of sigma² values
sigma2_vals <- seq(0.01, 500, length.out = 300)

# Closed-form probability: P(||X||^2 <= r^2)
bull_probs <- pchisq((r_bull^2) / sigma2_vals, df = 2)

# Plot
plot(sigma2_vals, bull_probs, type = "l", lwd = 2,
     xlab = expression(sigma^2), ylab = "Probability of Bullseye",
     main = "P(Landing in Bullseye) vs. " ~ sigma^2)
abline(h = 1, col = "gray", lty = 2)

```

### LogNormal Distribution stuff

```{r}
# Set seed for reproducibility (optional)
set.seed(123)

# Number of players (or samples)
n_players <- 5

# Sample sigma^2 from LogNormal(meanlog = 1, sdlog = 0.25)
sigma2_true <- rlnorm(n_players, meanlog = 2, sdlog = 0.25)

# Print the sampled values
print(sigma2_true)
```
